---
title: "matrix_test"
author: "jcrull2"
date: "2023-04-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Import data and convert to realRatingMatrix

```{r}
#install.packages("recommenderlab")
library(tidyverse)
library(recommenderlab)
library(softImpute)
library()
library()
my_matrix <- data.matrix(subset(read.csv('feedback.csv'),select=c(-1)), rownames.force = NA)
paste0('Number of nonmissing ratings: ',length(my_matrix) - length(which(is.na(my_matrix))))
paste0('Number of missing ratings: ',length(which(is.na(my_matrix))))
real_matrix = as(my_matrix,"realRatingMatrix")
real_matrix
```
### EDA

```{r}
getRatingMatrix(real_matrix)
```

```{r}
summary(getRatings(real_matrix))
data.frame(ratings = getRatings(real_matrix)) %>%
  ggplot(aes(ratings)) + geom_bar(width = 0.75) +
    labs(title = 'Restaurant Ratings Distribution')

data.frame(rowcount = rowCounts(real_matrix)) %>%
  ggplot(aes(rowcount/ncol(real_matrix))) + geom_bar(width = 0.05) +
    labs(title = 'Proportion of Ratings by Respondent')

data.frame(colcount = colCounts(real_matrix)) %>%
  ggplot(aes(colcount/nrow(real_matrix))) + geom_bar(width = 0.02) +
    labs(title = 'Proportion of Ratings by Restaurant')
```

```{r}
print("Mean ratings by respondent")
rowMeans(real_matrix)

print("Mean ratings by restaurant")
colMeans(real_matrix)
```


```{r}
 recommenderRegistry$get_entries(dataType = "realRatingMatrix")
```

### Split data into train/test rows (80/20), giving three ratings per test row to algorithm and evaluate recommenders on remaining ratings from each test respondent

```{r}
set.seed(42)
print(paste0("Minimum number of ratings per respondent: ",min(rowCounts(real_matrix))))

#n_matrix=normalize(real_matrix)
#Algorithms normalize matrix by subtracting row means
#Split data into 80% train/20% test, evaluate on all but three per row in test
e <- evaluationScheme(real_matrix, method="split", train=0.8,given=3, goodRating=3)
r_random <- Recommender(getData(e, "train"), "RANDOM")
r_popular <- Recommender(getData(e, "train"), "POPULAR")
r_ubcf <- Recommender(getData(e, "train"), "UBCF")
r_ibcf <- Recommender(getData(e, "train"), "IBCF")
r_svdf <- Recommender(getData(e, "train"), "SVDF")

p_random <- predict(r_random, getData(e, "known"), type="ratings")
p_popular <- predict(r_popular, getData(e, "known"), type="ratings")
p_ubcf <- predict(r_ubcf, getData(e, "known"), type="ratings")
p_ibcf <- predict(r_ibcf, getData(e, "known"), type="ratings")
p_svdf <- predict(r_svdf, getData(e, "known"), type="ratings")

error <- rbind(RANDOM = calcPredictionAccuracy(p_random, getData(e, "unknown")),POPULAR = calcPredictionAccuracy(p_popular, getData(e, "unknown")),UBCF = calcPredictionAccuracy(p_ubcf, getData(e, "unknown")),IBCF = calcPredictionAccuracy(p_ibcf, getData(e, "unknown")),SVDF = calcPredictionAccuracy(p_svdf, getData(e, "unknown")))
error
paste0("RMSE with just the training mean: ",round((mean((getRatings(getData(e,"unknown"))-mean(getRatings(getData(e,"train"))))^2))^.5,3))
#getRatingMatrix(p_ubcf)
#getRatingMatrix(real_matrix)
#getRatingMatrix(getData(e,"unknown"))
#getRatingMatrix(getData(e,"known"))
```

### Somehow these algorithms perform worse than the naive estimator? Double-check calculations
### Future: explore parameter tuning (if applicable)


### softImpute (***Randomly setting 40 obs to NA for evaluation and training on rest***)
```{r}
print(paste0("Number of values used to calculate RMSE above: ",length(getRatings(getData(e,"unknown")))))
#softImpute
set.seed(42)
test_obs = sample(setdiff(1:length(my_matrix),which(is.na(my_matrix))),40,replace=FALSE)
test = my_matrix[test_obs]
train = my_matrix
train[test_obs] <- NA
nonmissing_mean = mean(getRatings(as(train,"realRatingMatrix")))


standardized_matrix = train-nonmissing_mean
fits_als <- softImpute(standardized_matrix, trace=FALSE, type = "als",lambda=0)
hardImpute_matrix_als = (fits_als$u %*% diag(fits_als$d) %*% t(fits_als$v)) + nonmissing_mean
fits_svd <- softImpute(standardized_matrix, trace=FALSE, type = "svd",lambda=0)
hardImpute_matrix_svd = (fits_svd$u %*% diag(fits_svd$d) %*% t(fits_svd$v)) + nonmissing_mean

#Tuning max.rank and lambda for ALS
best_rmse_als = 69
best_lambda_als = 0
best_max_rank_als = 0
for (max_rank in 2:14){  
  for (lambda in seq(0,8,0.1)){
    set.seed(42)
    fits <- softImpute(standardized_matrix, trace=FALSE, type = "als",lambda=lambda,rank.max=max_rank)
    new_matrix = (fits$u %*% diag(fits$d) %*% t(fits$v)) + nonmissing_mean
    rmse = (mean((test-new_matrix[test_obs])^2))^.5
    #print(paste0("RMSE with lambda=",lambda," and max rank = ",max_rank,": ",round(rmse)))
    if (rmse<best_rmse_als){
      best_rmse_als = rmse
      best_lambda_als = lambda
      best_max_rank_als = max_rank
    }
  }
}

#Tuning max.rank and lambda for SVD
best_rmse_svd = 69
best_lambda_svd = 0
best_max_rank_svd = 0
for (max_rank in 2:14){  
  for (lambda in seq(0,8,0.1)){
    set.seed(42)
    fits <- softImpute(standardized_matrix, trace=FALSE, type = "svd",lambda=lambda,rank.max=max_rank)
    new_matrix = (fits$u %*% diag(fits$d) %*% t(fits$v)) + nonmissing_mean
    rmse = (mean((test-new_matrix[test_obs])^2))^.5
    #print(paste0("RMSE with lambda=",lambda," and max rank = ",max_rank,": ",round(rmse)))
    if (rmse<best_rmse_svd){
      best_rmse_svd = rmse
      best_lambda_svd = lambda
      best_max_rank_svd = max_rank
    }
  }
}

paste0("RMSE with just the mean: ",round((mean((test-nonmissing_mean)^2))^.5,3))
print(paste0("RMSE with ALS hardImpute: ",round((mean((test-hardImpute_matrix_als[test_obs])^2))^.5,3)))
print(paste0("RMSE with SVD hardImpute: ",round((mean((test-hardImpute_matrix_svd[test_obs])^2))^.5,3)))

print(paste0("Best parameters for ALS: lambda=",best_lambda_als," ,max.rank=",best_max_rank_als))
print(paste0("RMSE with tuned softImpute with ALS: ",round(best_rmse_als,3)))

print(paste0("Best parameters for SVD: lambda=",best_lambda_svd," ,max.rank=",best_max_rank_svd))
print(paste0("RMSE with tuned softImpute with SVD: ",round(best_rmse_svd,3)))
```
### Worth noting again that these RMSEs may not be comparable to those of the above Recommenders because we use different observations to evaluate and train on a larger set

